Magnetized accretion disks around Kerr black holes with scalar hair - I. Constant angular momentum

### 2018-07-30 paper_prd.tex
fixed table I typos
Some work on the draft.

In addition, the detection of gravitational waves from binary BHs~\cite{Abbott2016, Abbott:2016nmj, Abbott:2017vtc, Abbott:2017oio, Abbott:2017gyy} and the prospect of observing the first image - the black hole {\it shadow} - of a BH by the Event Horizon Telescope~\cite{Fish:2016} opens the opportunity to test the true nature of BHs and the astrophysical relevance of HBHs. 
Regarding the latter, the source of light in a realistic situation is provided by material from an accretion disk crossing the event horizon of the BH, producing the shadow. Thick accretion disks (or tori) are common in astrophysics, either surrounding the central BHs of quasars and active galactic nuclei or, at a stellar scale, surrounding the compact objects in X-ray binaries, microquasars, and gamma-ray bursts. Motivated by this observational possibility, in this work we build equilibrium solutions of HBHs surrounded by magnetized accretion disks. Our procedure, which combines earlier approaches  
by~\cite{Komissarov:2006,Qian:2009} was presented in~\cite{Gimeno-Soler:2017} for the Kerr BH case. In~\cite{Gimeno-Soler:2017} we built equilibrium sequences of accretion disks  endowed with a purely toroidal magnetic field, assuming a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. Our immediate goal in the present work is to assess if and how the accretion disk morphology depends on the type of BH considered, either Kerr BHs of varying spins or Kerr BHs with scalar hair (KBHsSH). Moreover, we focus on disks with a constant distribution of specific angular momentum, presenting in a companion paper the non-constant (power-law) case. In subsequent investigations, we will analyze the shadows of the equilibrium models of KBHsSH presented in this work as well as in {\it dynamical} situations.

@@

%%%%%%%%%%%%%%%%%%%
\begin{table*}
\caption{List of models of KBHsSH used in this work. From left to right the columns report the name of the model, the ADM mass, $M_{\mathrm{ADM}}$, the ADM angular momentum, $J_{\mathrm{ADM}}$, the horizon mass, $M_{\mathrm{H}}$, the horizon angular momentum, $J_{\mathrm{H}}$, the mass of the scalar field, $M_{\mathrm{SF}}$, the angular momentum of the scalar field, $J_{\mathrm{SF}}$, the radius of the event horizon, $r_{\mathrm{H}}$, the values of the normalized spin parameter for the ADM quantities, $a_{\mathrm{ADM}}$, and for the BH horizon quantities, $a_{\mathrm{H}}$, the horizon linear velocity, $v_{\mathrm{H}}$, and the spin parameter corresponding to a Kerr BH with a linear velocity equal to $v_{\mathrm{H}}$, $a_{\mathrm{H_{eq}}}$.}        
\label{models_list}      
\centering          
\begin{tabular}{c c c c  c c c c   c c c c}
\hline\hline       
 Model & $M_{\mathrm{ADM}}$ & $J_{\mathrm{ADM}}$ & $M_{\mathrm{H}}$ &  $J_{\mathrm{H}}$ & $M_{\mathrm{SF}}$ & $J_{\mathrm{SF}}$ & $r_{\mathrm{H}}$ & $a_{\mathrm{ADM}}$ & $a_{\mathrm{H}}$ & $v_{\mathrm{H}}$ & $a_{\mathrm{H_{eq}}}$ \\ 
\hline           
I & $0.415$ & $0.172$ & $0.393$ &  $0.15$  & $0.022$ & $0.022$ & $0.200$ & $0.9987$ & $0.971$ & $0.7685$ & $0.9663$\\ 
 \hline 
II & $0.630$ & $0.403$ & $0.340$ &  $0.121$  & $0.063$ & $0.282$ & $0.221$ & $1.0140$ & $0.376$ & $0.6802$ & $0.9301$ \\
 \hline 
III & $0.797$ & $0.573$ & $0.365$ &  $0.172$  & $0.573$ & $0.432$ & $0.111$ & $0.9032$ & $1.295$ & $0.7524$ & $0.9608$ \\ 
 \hline 
IV & $0.933$ & $0.739$ & $0.234$ &  $0.114$  & $0.699$ & $0.625$ & $0.100$ & $0.8489$ & $2.082$ & $0.5635$ & $0.8554$ \\ 
 \hline 
V & $0.940$ & $0.757$ & $0.159$ &  $0.076$  & $0.757$ & $0.781$ & $0.091$ & $0.8560$ & $3.017$ & $0.4438$ & $0.7415$ \\ 
 \hline 
VI & $0.959$ & $0.795$ & $0.087$ &  $0.034$  & $0.872$ & $0.781$ & $0.088$ & $0.9477$ & $3.947$ & $0.2988$ & $0.5487$ \\ 
 \hline 
VII & $0.975$ & $0.850$ & $0.018$ &  $0.002$  & $0.957$ & $0.848$ & $0.040$ & $0.8941$ & $6.173$ & $0.0973$ & $0.1928$ \\ 
\hline      
\end{tabular}
\end{table*}
%%%%%%%%%%%%%%%%%%%

@@

The stationary and axisymmetric metric ansatz is a solution to the EKG field equations $R_{ab} - \frac{1}{2}R g _{ab} = 8 \pi G (T_{\mathrm{SF}})_{ab}$ with 
\begin{eqnarray}\label{eq:e-m_scalaf_field}
(T_{\mathrm{SF}})_{ab} &=& \partial_a \Psi^* \partial_b \Psi + \partial_b \Psi^* \partial_a \Psi 
\nonumber \\ 
&-& g_{ab} \left(\frac{1}{2} g^{cd}(\partial_c \Psi^* \partial_d \Psi + \partial_d \Psi^* \partial_c \Psi) + \mu^2 \Psi^* \Psi \right),
\end{eqnarray}
where $\mu$ is the mass of the scalar field and $^*$ denotes complex conjugation. The interested reader is addressed to~\cite{Herdeiro:2015b} for details on the equations of motion for the scalar field $\Psi$ and the four metric functions ${W, F_i}$, along with their solution.

Table~\ref{models_list} lists the seven KBHsSH models we use in this work. The models have been selected to span all regions of interest in the parameter space. Model I corresponds to a Kerr-like model, with almost all the mass and angular momentum stored in the BH, while model VII corresponds to a KBHSH with almost all the mass and angular momentum stored in the scalar field. It is worth mentioning that some of the models violate the Kerr bound (normalized sìn parameter larger than 1) in terms of both ADM or horizon quantities. This is not a source of concern because, as shown in~\cite{Herdeiro:2015c} the linear velocity of the horizon, $v_{\mathrm{H}}$, never exceeds $1$. For comparison, we also show in the last column of Table~\ref{models_list} the spin parameter $a_{\mathrm{H_{eq}}}$ corresponding to a Kerr BH with a horizon linear velocity $v_{\mathrm{H}}$. \tf{Further description of the table. Something must be said about the coordinates.}. In addition to the information provided in Table~\ref{models_list}, Figure~\ref{existence} plots the location of our models in  the domain of existence of KBHsSH in an ADM mass versus scalar field frequency diagram.

@@

Equilibrium models of disks around Kerr BHs are built assuming that the spacetime metric and the fluid fields are stationary and  axisymmetric.  This was the approach followed in~\cite{Gimeno-Soler:2017}. For KBHsSH we can follow the same approach as the metric ansatz given by Eq.~(\ref{metric}) is stationary and axisymmetric.

@@

In this work, for simplicity, we consider a constant angular momentum distribution, $l(r,\theta) = \mathrm{const}$. This is contrary to the approach we followed in~\cite{Gimeno-Soler:2017} where the angular momentum ansatz followed the approach of~\cite{Qian:2009}. \tf{We should explain/motivate why we change our approach in this paper.} The specific value of the angular momentum is computed as the minimum of the following equation
\begin{equation}\label{eq:mb_ang_mom}
l^{\pm}_{\mathrm{b}}(r, \theta) = \frac{g_{t\phi}\pm\left(\sqrt{g_{t\phi}^2-g_{tt}g_{\phi\phi}}\right)\sqrt{1+g_{tt}}}{-g_{tt}}
\end{equation}
where the plus sign is for prograde orbits and the minus is for retrograde orbits.
This expression is given by~\citep{Daigne:2004} for Kerr BHs, but it is valid for any stationary, axisymmetric spacetimes. For prograde motion, the function has a minimum outside the event horizon. The location of said minimum corresponds with the marginally bound orbit $r_{\mathrm{mb}}$, and the angular momentum corresponds to the keplerian angular momentum $l_{\mathrm{mb}}$ at that point. We show the proof of this statement at appendix~\ref{ang_mom_appendix}. This choice of angular momentum distribution is motivated by its simplicity (for a first study of thick tori around KBHsSH) and for allowing the presence of a cusp (to allow matter accretion onto the black hole) and a centre.

@@

We use the procedure described by~\cite{Montero:2007}, where we write the equations of ideal general relativistic MHD as the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (\rho h + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $h$, $\rho$ $p$ being the fluid specific enthalpy, density and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting the divergence of Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(\rho h + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Then, we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.

To integrate Eq.~\eqref{eq:diff_ver} we first assume a polytropic equation of state of the form
\begin{equation}\label{eq:eos_fluid}
p = K \rho^{\Gamma},
\end{equation}
with $K$ and $\Gamma$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, the magnetization parameter $\beta_{\mathrm{m}} = p/p_{\mathrm{m}}$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogous equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = M \tilde{w}^q,
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = M \mathcal{L}^{q-1} w^q,
\end{equation}
where $w = \rho h$ is the fluid enthaply density, and $M$ and $q$ are constants.
If we define the potential as $W \equiv \ln |u_t|$, then we can integrate the equation~\eqref{eq:diff_ver} as
\begin{equation}\label{eq:final}
W - W_{\mathrm{in}} + \ln \left(1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}\right) + \frac{q}{q-1}M(\mathcal{L}w)^{q-1}=0,
\end{equation}
where $W_{\mathrm{in}}$ is the potential at the inner edge of the disk.

We also write the expressions of the total energy density for the torus $\rho_{\mathrm{T}}$:
\begin{equation}\label{eq:torus_energy_density}
\rho_{\mathrm{T}} = -T^t_t + T^i_i = \frac{\rho h (g_{\phi\phi} - g_{tt} l^2)}{g_{\phi\phi} + 2 g_{t\phi} l + g_{tt} l^2} + 2 (p + p_{\mathrm{m}}),
\end{equation}
and the total energy density for the scalar field $\rho_{\mathrm{SF}}$:
\begin{equation}\label{eq:SF_energy_density}
\rho_{\mathrm{SF}} = -(T_{\mathrm{SF}})^t_t + (T_{\mathrm{SF}})^i_i = 2 \left(\frac{2 e^{2 F_0} w (w-m W)}{N} - \mu^2\right) \phi^2.
\end{equation}

In this work, we set the mass of the scalar field $\mu = 1$, the angular momentum to $l = l_{\mathrm{mb}}$, the inner radius of the disk to $r_{\mathrm{in}} = r_{\mathrm{mb}}$, the exponents of the polytropic EoS to $q = \Gamma = 4/3$, and the density at the disk centre $\rho_{\mathrm{c}} = 1$. Then, we leave the magnetization at the center $\beta_{\mathrm{m_c}}$ as a parameter (for each model of KBHSH). With this information we can compute all the relevant physical quantities.

@@

%%%%%%%%%%%%%%%%
\subsection{Building the disk}
%%%%%%%%%%%%%%%%

To construct our models we take the following steps:
First, we find the angular momentum $l$ and the radius of the cusp $r_{\mathrm{cusp}}$ as the value at the minimum and the location of said minimum outside the event horizon of equation~\eqref{eq:mb_ang_mom}. This choice of angular momentum implies $r_{\mathrm{cusp}} = r_{\mathrm{mb}}$, $l = l_{\mathrm{K}}(r_{\mathrm{mb}})$ and $W_{\mathrm{in}} = 0$ and then, we can compute the potential distribution as $W(r, \theta) \equiv \ln |u_t|$. Considering this, and that we can write the magnetic pressure at the center as
\begin{equation}
p_{\mathrm{m_c}} = M \mathcal{L}^{q-1} \left(1 + \frac{\Gamma K}{\Gamma +1}\right)^q = K/\beta_{\mathrm{m_c}},
\end{equation}
where we have taken into account that $\rho_{\mathrm{c}} = 1$, that the pressure at the center is $p_{\mathrm{c}} = K$ and the specific enthalpy can be written as $h = 1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}$. Replacing this into the equation~\eqref{eq:final} we arrive at an equation for the polytropic constant $K$
\begin{equation}\label{eq:K_eq}
W_{\mathrm{c}} + \ln \left(1 + \frac{\Gamma K}{\Gamma +1}\right) + \frac{q}{q-1}\frac{K/\beta_{\mathrm{m_c}}}{1 + \frac{\Gamma K}{\Gamma +1}} = 0.
\end{equation}
Once we have $K$, it is easy to find $p_{\mathrm{m}}$, $M$ and $h_{\mathrm{c}}$ with the definition of the magnetization parameter at the center $\beta_{\mathrm{m_c}}$, the equation~\eqref{eq:eos_mag} and $h = 1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}$. With this, we already have all we need in order to compute all the relevant physical quantities inside the disk.

>>>>

In addition, the detection of gravitational waves from binary BHs~\cite{Abbott2016, Abbott:2016nmj, Abbott:2017vtc, Abbott:2017oio, Abbott:2017gyy} and the prospect of observing the first image - the black hole {\it shadow} - of a BH by the Event Horizon Telescope~\cite{Fish:2016} opens the opportunity to test the true nature of BHs and the astrophysical relevance of HBHs. 
Regarding the latter, the source of light in a realistic situation is provided by material from an accretion disk crossing the event horizon of the BH, producing the shadow. Thick accretion disks (or tori) are common systems in astrophysics, either surrounding the central BHs of quasars and active galactic nuclei or, at a stellar scale, surrounding the compact objects in X-ray binaries, microquasars, and gamma-ray bursts. Motivated by this observational possibility, in this work we build equilibrium solutions of HBHs surrounded by magnetized accretion disks. Our procedure, which combines earlier approaches  
by~\cite{Komissarov:2006,Qian:2009} was presented in~\cite{Gimeno-Soler:2017} for the Kerr BH case. In~\cite{Gimeno-Soler:2017} we built equilibrium sequences of accretion disks  endowed with a purely toroidal magnetic field, assuming a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. Our immediate goal in the present work is to assess if and how the accretion disk morphology depends on the type of BH considered, either Kerr BHs of varying spins or Kerr BHs with scalar hair (KBHsSH). Moreover, we focus on disks with a constant distribution of specific angular momentum, presenting in a companion paper the non-constant (power-law) case. In subsequent investigations, we will analyze the shadows of the equilibrium models of KBHsSH presented in this work as well as in {\it dynamical} situations.

@@

%%%%%%%%%%%%%%%%%%%
\begin{table*}
\caption{List of models of KBHsSH used in this work. From left to right the columns report the name of the model, the ADM mass, $M_{\mathrm{ADM}}$, the ADM angular momentum, $J_{\mathrm{ADM}}$, the horizon mass, $M_{\mathrm{H}}$, the horizon angular momentum, $J_{\mathrm{H}}$, the mass of the scalar field, $M_{\mathrm{SF}}$, the angular momentum of the scalar field, $J_{\mathrm{SF}}$, the radius of the event horizon, $r_{\mathrm{H}}$, the values of the normalized spin parameter for the ADM quantities, $a_{\mathrm{ADM}}$, and for the BH horizon quantities, $a_{\mathrm{H}}$, the horizon linear velocity, $v_{\mathrm{H}}$, and the spin parameter corresponding to a Kerr BH with a linear velocity equal to $v_{\mathrm{H}}$, $a_{\mathrm{H_{eq}}}$.}        
\label{models_list}      
\centering          
\begin{tabular}{c c c c  c c c c   c c c c}
\hline\hline       
 Model & $M_{\mathrm{ADM}}$ & $J_{\mathrm{ADM}}$ & $M_{\mathrm{H}}$ &  $J_{\mathrm{H}}$ & $M_{\mathrm{SF}}$ & $J_{\mathrm{SF}}$ & $r_{\mathrm{H}}$ & $a_{\mathrm{ADM}}$ & $a_{\mathrm{H}}$ & $v_{\mathrm{H}}$ & $a_{\mathrm{H_{eq}}}$ \\ 
\hline           
I & $0.415$ & $0.172$ & $0.393$ &  $0.15$  & $0.022$ & $0.022$ & $0.200$ & $0.9987$ & $0.971$ & $0.7685$ & $0.9663$\\ 
 \hline 
II & $0.630$ & $0.403$ & $0.340$ &  $0.121$  & $0.290$ & $0.282$ & $0.221$ & $1.0140$ & $0.376$ & $0.6802$ & $0.9301$ \\
 \hline 
III & $0.797$ & $0.573$ & $0.365$ &  $0.172$  & $0.432$ & $0.401$ & $0.111$ & $0.9032$ & $1.295$ & $0.7524$ & $0.9608$ \\ 
 \hline 
IV & $0.933$ & $0.739$ & $0.234$ &  $0.114$  & $0.699$ & $0.625$ & $0.100$ & $0.8489$ & $2.082$ & $0.5635$ & $0.8554$ \\ 
 \hline 
V & $0.940$ & $0.757$ & $0.159$ &  $0.076$  & $0.781$ & $0.680$ & $0.091$ & $0.8560$ & $3.017$ & $0.4438$ & $0.7415$ \\ 
 \hline 
VI & $0.959$ & $0.795$ & $0.087$ &  $0.034$  & $0.781$ & $0.747$ & $0.088$ & $0.9477$ & $3.947$ & $0.2988$ & $0.5487$ \\ 
 \hline 
VII & $0.975$ & $0.850$ & $0.018$ &  $0.002$  & $0.957$ & $0.848$ & $0.040$ & $0.8941$ & $6.173$ & $0.0973$ & $0.1928$ \\ 
\hline      
\end{tabular}
\end{table*}
%%%%%%%%%%%%%%%%%%%

@@

The stationary and axisymmetric metric ansatz is a solution to the EKG field equations $R_{ab} - \frac{1}{2}R g _{ab} = 8 \pi G (T_{\mathrm{SF}})_{ab}$ with 
\begin{eqnarray}\label{eq:e-m_scalaf_field}
(T_{\mathrm{SF}})_{ab} &=& \partial_a \Psi^* \partial_b \Psi + \partial_b \Psi^* \partial_a \Psi 
\nonumber \\ 
&-& g_{ab} \left(\frac{1}{2} g^{cd}(\partial_c \Psi^* \partial_d \Psi + \partial_d \Psi^* \partial_c \Psi) + \mu^2 \Psi^* \Psi \right),
\end{eqnarray}
where $\mu$ is the mass of the scalar field and superscript $(^*)$ denotes complex conjugation. The interested reader is addressed to~\cite{Herdeiro:2015b} for details on the equations of motion for the scalar field $\Psi$ and the four metric functions ${W, F_0,F_1,F_2}$, along with their solution.

Table~\ref{models_list} lists the seven KBHsSH models we use in this work. The models have been selected to span all regions of interest in the parameter space. Model I corresponds to a Kerr-like model, with almost all the mass and angular momentum stored in the BH, while model VII corresponds to a KBHSH with almost all the mass and angular momentum stored in the scalar field. It is worth mentioning that some of the models violate the Kerr bound (normalized spìn parameter larger than 1) in terms of both ADM or horizon quantities. This is not a source of concern because, as shown in~\cite{Herdeiro:2015c} the linear velocity of the horizon, $v_{\mathrm{H}}$, never exceeds $1$. For comparison, we also show in the last column of Table~\ref{models_list} the spin parameter $a_{\mathrm{H_{eq}}}$ corresponding to a Kerr BH with a horizon linear velocity $v_{\mathrm{H}}$. \tf{Further description of the table. Something must be said about the coordinates.}. In addition to the information provided in Table~\ref{models_list}, Figure~\ref{existence} plots the location of our models in  the domain of existence of KBHsSH in an ADM mass versus scalar field frequency diagram.

@@

Equilibrium models of thick disks around Kerr BHs are built assuming that the spacetime metric and the fluid fields are stationary and  axisymmetric (see, e.g.~\cite{Font:2002,Daigne:2004} and references therein).  This was the approach followed in~\cite{Gimeno-Soler:2017}. For disks around KBHsSH we can follow the same approach as the metric ansatz given by Eq.~(\ref{metric}) is stationary and axisymmetric.

@@

The approach we followed in~\cite{Gimeno-Soler:2017} for the angular momentum distribution of the disks was introduced by~\cite{Qian:2009}, and it is characterized by three free parameters, $\beta$, $\gamma$, and $\eta$ (see Eq.~(7) in~\cite{Gimeno-Soler:2017}). In this work, for simplicity and to reduce the ample space of parameters, we consider a constant angular momentum distribution, $l(r,\theta) = \mathrm{const}$, which corresponds to setting $\beta=\gamma=0$ in~\cite{Gimeno-Soler:2017}. This choice also allows for the presence of a cusp (and hence matter accretion onto the black hole) and a centre.  \tf{We should explain/motivate why we change our approach in this paper.} The specific value of the angular momentum is computed as the minimum of the following equation
\begin{equation}\label{eq:mb_ang_mom}
l^{\pm}_{\mathrm{b}}(r, \theta) = \frac{g_{t\phi} \pm\left(\sqrt{g_{t\phi}^2-g_{tt}g_{\phi\phi}}\right)\sqrt{1+g_{tt}}}{-g_{tt}}
\end{equation}
where the plus sign corresponds to prograde orbits and the minus sign to retrograde orbits.
This expression is given by~\citep{Daigne:2004} for Kerr BHs, but it is valid for any stationary and axisymmetric spacetime. For prograde motion, the function has a minimum outside the event horizon. The location of this minimum corresponds with the marginally bound orbit $r_{\mathrm{mb}}$, and the angular momentum corresponds to the Keplerian angular momentum $l_{\mathrm{mb}}$ at that point. We show the proof of this statement in Appendix~\ref{ang_mom_appendix}. \tf{Is there an Appendix A?}.

@@

To account for the magnetic flied in the disks we use the procedure described by~\cite{Montero:2007}. We write the equations of ideal general relativistic MHD as the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (\rho h + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetized perfect fluid, $h$, $\rho$, and  $p$ being the fluid specific enthalpy, density and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$ (see~\cite{Anton:2006} for further details). Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting the divergence of Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(\rho h + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Then, we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.

To integrate Eq.~\eqref{eq:diff_ver} we first assume a polytropic equation of state of the form
\begin{equation}\label{eq:eos_fluid}
p = K \rho^{\Gamma},
\end{equation}
with $K$ and $\Gamma$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, the magnetization parameter $\beta_{\mathrm{m}} = p/p_{\mathrm{m}}$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogous equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = M \tilde{w}^q,
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = M \mathcal{L}^{q-1} w^q,
\end{equation}
where $w = \rho h$ is the fluid enthaply density, and $M$ and $q$ are constants.
If we define the potential as $W \equiv \ln |u_t|$, then we can integrate Eq.~\eqref{eq:diff_ver} as
\begin{equation}\label{eq:final}
W - W_{\mathrm{in}} + \ln \left(1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}\right) + \frac{q}{q-1}M(\mathcal{L}w)^{q-1}=0,
\end{equation}
where $W_{\mathrm{in}}$ is the potential at the inner edge of the disk.

In addition, the total energy density for the torus $\rho_{\mathrm{T}}$ is given by
\begin{equation}\label{eq:torus_energy_density}
\rho_{\mathrm{T}} = -T^t_t + T^i_i = \frac{\rho h (g_{\phi\phi} - g_{tt} l^2)}{g_{\phi\phi} + 2 g_{t\phi} l + g_{tt} l^2} + 2 (p + p_{\mathrm{m}}),
\end{equation}
while the total energy density for the scalar field $\rho_{\mathrm{SF}}$ reads
\begin{equation}\label{eq:SF_energy_density}
\rho_{\mathrm{SF}} = -(T_{\mathrm{SF}})^t_t + (T_{\mathrm{SF}})^i_i = 2 \left(\frac{2 e^{2 F_0} w (w-m W)}{N} - \mu^2\right) \phi^2.
\end{equation}

In this work, we set the mass of the scalar field $\mu = 1$, the angular momentum to $l = l_{\mathrm{mb}}$, the inner radius of the disk to $r_{\mathrm{in}} = r_{\mathrm{mb}}$, the exponents of the polytropic EoS to $q = \Gamma = 4/3$, and the density at the centre of the disk $\rho_{\mathrm{c}} = 1$. Thus, we leave the magnetization at the center $\beta_{\mathrm{m_c}}$ as a free parameter for each model of KBHSH. With this information we can compute all relevant physical quantities.

@@

%%%%%%%%%%%%%%%%
\subsection{Building the disks}
%%%%%%%%%%%%%%%%

To construct our models we take the following steps:
First, we find the angular momentum $l$ and the radius of the cusp $r_{\mathrm{cusp}}$ as the value at the minimum of Eq.~\eqref{eq:mb_ang_mom} and the location of said minimum outside the event horizon, respectively. This choice of angular momentum implies $r_{\mathrm{cusp}} = r_{\mathrm{mb}}$, $l = l_{\mathrm{K}}(r_{\mathrm{mb}})$, and $W_{\mathrm{in}} = 0$. Next, we can compute the gravitational potential distribution as $W(r, \theta) \equiv \ln |u_t|$. Considering this, and that we can write the magnetic pressure at the centre as
\begin{equation}
p_{\mathrm{m_c}} = M \mathcal{L}^{q-1} \left(1 + \frac{\Gamma K}{\Gamma +1}\right)^q = \frac{K}{\beta_{\mathrm{m_c}}},
\end{equation}
where we have taken into account that $\rho_{\mathrm{c}} = 1$, that the pressure at the center is $p_{\mathrm{c}} = K$ and the specific enthalpy can be written as $h = 1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}$. \tf{The previous sentence reads a bit strange.} Replacing this into Eq.~\eqref{eq:final} we arrive at an equation for the polytropic constant $K$
\begin{equation}\label{eq:K_eq}
W_{\mathrm{c}} + \ln \left(1 + \frac{\Gamma K}{\Gamma +1}\right) + \frac{q}{q-1}\frac{K/\beta_{\mathrm{m_c}}}{1 + \frac{\Gamma K}{\Gamma +1}} = 0.
\end{equation}
Once we have $K$, it is easy to find $p_{\mathrm{m}}$, $M$ and $h_{\mathrm{c}}$ with the definition of the magnetization parameter at the center $\beta_{\mathrm{m_c}}$, the equation~\eqref{eq:eos_mag} and $h = 1 + \frac{\Gamma K}{\Gamma +1}\rho^{\Gamma -1}$. With this, we already have all we need in order to compute all the relevant physical quantities inside the disk.
